---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE}
        library(readr)
        activity <- read_csv("~/RepData_PeerAssessment1/activity.csv")
```


## What is mean total number of steps taken per day?
* First, make a histogram of the total number of steps taken each day:

```{r echo = TRUE}
        library(ggplot2)
        steps_per_day <- aggregate(activity$steps, list(activity$date), sum)
        qplot(activity$steps, geom="histogram") 
```

* Second, calculate and report the mean and median total number of steps taken per day:

```{r echo = TRUE}
       mean_steps <- mean(steps_per_day$x, na.rm = TRUE)
```
```{r echo = TRUE}
       median_steps <- median(steps_per_day$x, na.rm = TRUE)
```
Mean number of steps per day is `r mean_steps`, while median number of steps per day is `r median_steps`

## What is the average daily activity pattern?

```{r echo = TRUE}
steps_per_interval <- aggregate(activity$steps, list(activity$interval), function(x) {mean(x, na.rm = TRUE)})
ggplot(steps_per_interval, aes(strptime(Group.1, format="%H%M"), x)) + geom_line()+ xlab("Time") + ylab("Number of Steps")
```

```{r echo = TRUE}
max_steps <- max(steps_per_interval$x)
max_interval <- steps_per_interval[steps_per_interval$x == max_steps,]$Group.1
# max_interval <- strptime(max_interval, format="%H%M")
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
